---
title: "project5"
format: html
editor: visual
author: Blair Kim
date: April 30, 2025
bibliography: references.bib
execute: 
  warning: false
  message: false
---

```{r}
library(ggplot2)
library(sqldf)
library(tidyverse)
```

```{r}

con_traffic <- DBI::dbConnect(

  RMariaDB::MariaDB(),

  dbname = "traffic",

  host = Sys.getenv("TRAFFIC_HOST"),

  user = Sys.getenv("TRAFFIC_USER"),

  password = Sys.getenv("TRAFFIC_PWD")

)
```

```{r}
con_traffic <- DBI::dbConnect(
  RMariaDB::MariaDB(),
  dbname = "traffic",
  host = Sys.getenv("TRAFFIC_HOST"),
  user = Sys.getenv("TRAFFIC_USER"),
  password = Sys.getenv("TRAFFIC_PWD")
)
```

```{sql connection=con_traffic}
SHOW TABLES;
```

Number of vehicle stops by race in big cities in California - San Francisco, Los Angeles, and San Diego

```{sql connection=con_traffic}
#| output.var: "sf_vehicle_stops"

select count(*) as vehicle_stops,
subject_race as race
from ca_san_francisco_2020_04_01	
where type = 'vehicular'
group by subject_race
order by vehicle_stops DESC
```

```{r}
sf_vehicle_stops 
```

```{sql connection=con_traffic}
#| output.var: "la_vehicle_stops"

select count(*) as vehicle_stops,
subject_race as race
from ca_los_angeles_2020_04_01		
where type = 'vehicular'
group by subject_race
order by vehicle_stops DESC

```

```{r}
la_vehicle_stops 
```

```{sql connection=con_traffic}
#| output.var: "sd_vehicle_stops"

select count(*) as vehicle_stops,
subject_race as race
from ca_san_diego_2020_04_01		
where type = 'vehicular'
group by subject_race
order by vehicle_stops DESC
```

```{r}
sd_vehicle_stops
```

Were more POC drivers stopped by police in the past then now in LA?

```{sql connection=con_traffic}
#| output.var: "la_POC_stops_by_year"

select subject_race as race, year(date) as year, count(*) as num_stops
from ca_los_angeles_2020_04_01
where subject_race != 'white' and date IS NOT NULL 
group by race, year(date)
order by year asc
```

```{r}
la_POC_stops_by_year
```

```{r}
# visualize POC drivers stops by year

la_POC_stops_by_year|> 
  mutate(num_stops = as.numeric(num_stops)) |> 
  ggplot(aes(x = year, y = num_stops, color = race)) +
  geom_line() +
  theme_minimal() +
  labs(title = "# of Police Stops of POC drivers in LA by year", y = "number of stops", x = "year") 

```

Among all racial groups, Hispanic drivers were most likely to be stopped by police in LA. It seems like there was a sharp increase in stops of POC drivers in 2011, but the numbers gradually decreased over time.

```{sql connection=con_traffic}
DESCRIBE ca_san_francisco_2020_04_01	
```

```{sql connection=con_traffic}
DESCRIBE ca_san_jose_2020_04_01	
```

What time of the day are people most likely to be arrested in San Jose?

```{sql connection=con_traffic}
#| output.var: "san_jose_arrests_by_time"
select hour(time) as hour, count(*) as cases 
from ca_san_jose_2020_04_01	
where outcome = 'arrest'  
group by hour(time)
order by cases desc 
```

```{r}
# visualize number of arrests by hour in San Jose
san_jose_arrests_by_time |>
  ggplot(aes(x = hour, y = cases)) +
  geom_line() +
  theme_minimal() +
  labs(title = "# of arrests by hour in San Jose", y = "number of arrests", x = "hour") 
```

What about in San Francisco?

```{sql connection=con_traffic}
#| output.var: "sf_arrests_by_time"
select hour(time) as hour, count(*) as cases 
from ca_san_francisco_2020_04_01	
where outcome = 'arrest'  
group by hour(time)
order by cases desc 
```

```{r}
# visualize number of arrests by hour in San Francisco
sf_arrests_by_time |>
  ggplot(aes(x = hour, y = cases)) +
  geom_line() +
  theme_minimal() +
  labs(title = "# of arrests by hour in San Francisco", y = "number of arrests", x = "hour") 
```

In San Jose, around 9 AM is the peak time for getting arrested on the road. In San Francisco, around 5pm and midnight is when people get arrested most often.

References

Pierson, Emma, Camelia Simoiu, Jan Overgoor, Sam Corbett-Davies, Daniel Jenson, Amy Shoemaker, Vignesh Ramachandran, et al. 2020. “A Large-Scale Analysis of Racial Disparities in Police Stops Across the United States.” *Nature Human Behaviour*, 1–10.
